Atmospheric escape and star-planet-interaction in the HAT-P-67 b system

### 2022-02-25 paper/paper1/ms.tex
Save a backup version of the paper

    The Helium Exospheres program with the Habitable Zone Planet Finder (HPF) spectrograph on the Hobby Eberly Telescope (HET) in West Texas has targeted 23 exoplanet systems over two years. Here, we present a study of HAT-P-67, an F subgiant hosting a low density planet with 0.3 $M_{\mathrm{Jup}}$ and 2.1 $R_{\mathrm{Jup}}$ on a close-in 4.8 day orbit. The high energy radiation environment should provide a uniquely harsh testing ground for examining atmospheric escape.  The planet shows prominent in-transit absorption in the He 10830 \AA triplet and infrared Ca II 8542 \AA triplet across 20 visits over the course of a year including in- and out-of-transit spectroscopy. We identify the rotational period of HAT-P-67 to be 5.417 days by using contemporaneous data from the Transiting Exoplanet Survey Satellite (TESS). We consider multiple interpretations of the HPF and TESS observations.  First, the period measured by TESS may be the rotational period of the star and thus, the star and planet are nearly-tidally locked. On the other hand, the period measured by TESS may result from SPI-induced hot spots on the surface that follow the magnetic field lines linked to the planet's orbital period. Finally, we consider the possibility of stellar surface inhomogeneity contaminating the He and Ca II transit signals.
\end{abstract}

@@

Exoplanets must possess a range of atmospheric properties well beyond what we have seen in our own Solar System.  Extreme radiation effects, atmospheric erosion, and unexpected chemistry must exist, but collecting examples of these phenomena remains at the edge of the possible today. The minuscule signal strengths of atmospheric shells has hidden all but the most conspicuous patterns of lines.  Helium 10830 $\AA$ has emerged as the most conspicuous of these, enabling ground-based detection with high-resolution near-infrared (IR) spectrographs.

A total of \censor{xx} systems have Helium 10830 exosphere measurements to date: observations of the most extreme stellar systems reveal dramatic levels of mass loss, up to \censor{xx} $\%/\mathrm{Gyr}$.

\citet{2020ApJ...894...97N} discusses the discovery of helium on exoplanets.

>>>>

    The Helium Exospheres program with the Habitable Zone Planet Finder (HPF) spectrograph on the Hobby Eberly Telescope (HET) in West Texas has targeted 23 exoplanet systems over two years. Here, we present a study of HAT-P-67, an F subgiant hosting a low density planet with 0.3 $M_{\mathrm{Jup}}$ and 2.1 $R_{\mathrm{Jup}}$ on a close-in 4.8 day orbit. The high energy radiation environment should provide a uniquely harsh testing ground for examining atmospheric escape. The planet shows prominent in-transit absorption in the He 10830 \AA~ triplet and infrared Ca II 8542 \AA~ triplet across 20 visits over the course of a year including in- and out-of-transit spectroscopy. We identify the rotational period of HAT-P-67 to be 5.417 days by using contemporaneous data from the Transiting Exoplanet Survey Satellite (TESS). We consider multiple interpretations of the HPF and TESS observations. First, the period measured by TESS may be the rotational period of the star and thus, the star and planet are nearly-tidally locked. On the other hand, the period measured by TESS may result from SPI-induced hot spots on the surface that follow the magnetic field lines linked to the planet's orbital period. Finally, we consider the possibility of stellar surface inhomogeneity contaminating the He and Ca II transit signals.
\end{abstract}

@@

Exoplanets must possess a dynamic range of atmospheric properties far beyond what we have seen in our own Solar System.  Extreme radiation effects, atmospheric erosion, and unexpected chemistry must exist, but collecting examples of these phenomena remains at the edge of the possible today. The minuscule signal strengths of atmospheric shells has hidden all but the most conspicuous phenomena.  Helium 10830 \AA has emerged as the most conspicuous exosphere indicator, enabling ground-based detection with high-resolution near-infrared (IR) spectrographs.

A total of \censor{xx} systems have Helium 10830 exosphere measurements to date: \object[GJ 3470]{GJ3470} \citep{2020ApJ...894...97N}, \object{HD 63433 c} \citep{2022AJ....163...68Z}, WASP 107 b \citep{2019A&A...623A..58A}.
